---
title: Baltimore County Permits
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
```

```{r}
library(tidyverse)

# Load data
permits <- read_csv("Permits.csv")

# Check structure of APPL_DATE
str(permits$APPL_DATE)

# View some unique values
unique(permits$APPL_DATE)

library(lubridate)

permits <- permits %>%
  mutate(appl_date = str_trim(as.character(APPL_DATE)),  # Remove leading/trailing spaces
         appl_date = gsub("\\+.*", "", appl_date),  # Remove timezone info (+00)
         appl_date = parse_date_time(appl_date, orders = c("ymd HMS", "ymd", "mdy HMS", "mdy")))  # Convert to date

# Check if conversion worked
head(permits$appl_date)

permits <- permits %>%
  mutate(year = year(appl_date)) %>%  # Extract year
  filter(zip %in% c("21219", "21222", "21224"))  # Filter for Dundalk ZIPs

# Group by year and count permits
permits_by_year <- permits %>%
  group_by(year) %>%
  summarise(count = n())

# View result
print(permits_by_year)




```{r}
str(permits$appl_date)
unique(permits$appl_date)
```

```{r}
library(tidyverse)
library(lubridate)

# Load data
permits <- read_csv("Permits.csv")

# Ensure ZIP codes are characters & clean dates
permits <- permits %>%
  mutate(
    zip = as.character(ZIP),  # Ensure ZIP is character
    appl_date = str_trim(as.character(APPL_DATE)),  # Remove extra spaces
    appl_date = gsub("\\+.*", "", appl_date),  # Remove timezone info
    appl_date = parse_date_time(appl_date, orders = c("ymd HMS", "ymd", "mdy HMS", "mdy")),  # Convert date
    year = year(appl_date)  # Extract year
  ) %>%
  filter(zip %in% c("21219", "21222", "21224"))  # Filter Dundalk ZIPs

# Group by year and count permits
permits_by_year <- permits %>%
  group_by(year) %>%
  summarise(count = n()) 

# View result
print(permits_by_year)
```

