---
title: "Parking Data"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

# Baltimore Parking Meter Data

Source: Fells Point info from pdf provided by Baltimore Parking
Authority 3/8 /2023 at 3:11 Mt Vernon info here:
<https://parking.baltimorecity.gov/demand-based-parking-meter-rate-setting/mt-vernon-demand-based-parking-meter-rate-setting>
Tiffany James from the parking authority is checking on Federal Hill
Harbor East info here:
<https://parking.baltimorecity.gov/harbor-east-parking-demand-parking-meter-rate-setting>
Central Downtown ingo here:
<https://parking.baltimorecity.gov/central-downtown-demand-based-parking-meter-rate-setting>

Parking facilities from 311 Open Baltimore:
<https://data.baltimorecity.gov/datasets/3d995dff7c424d39a59bb65d4af49486_0/explore?location=39.320514%2C-76.609286%2C11.58>
Residential Parking Permits from Open baltimore:

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(formattable)
library(janitor)
library(googlesheets4)
#install.packages("pdftools")
library(pdftools)

#install.packages("ggmap")
library(ggmap)
register_google(key = "XXXXX")

```

#Mt Vernon Import neighborhood, create name, address and clean

```{r}
googlesheets4::gs4_deauth()
mtvernon <- read_sheet("https://docs.google.com/spreadsheets/d/1FOL2vBfA4ifBbbjjjQEVDCHklcI_OWvd3vghDiBqDWE/edit#gid=1640964516", "Mt Vernon") %>% 
  as.data.frame() %>% 
  clean_names() %>% 
  mutate(neighborhood = "mtvernon") %>% 
  mutate(address = paste0(block, " ", street, ", ","Baltimore", ", ","Md")) %>% 
  mutate(initial_rate1 = as.numeric(unlist(initial_rate))) 

mtvernon <- mtvernon %>% 
  mutate(first_rate_change1 = as.numeric(as.character(first_rate_change))) %>% 
  mutate(second_rate_change1 = as.numeric(unlist(second_rate_change)))

#figure out why E Chase is .50 and .75; 800 N Howard is $1 and $2; 200 W Centre is $1 and $2; W Preston is .75 and $2 - what a huge range!

#need to geocode the streets


#Note on the brute force conversion of the first_rate_change1: see below

glimpse(mtvernon)
```

#analysis of mtvernon neighborhood

```{r}
mtvernon <- mtvernon %>% 
  mutate(first_diff = (first_rate_change1 - initial_rate1)) %>% 
  mutate(second_diff = (second_rate_change1 - first_rate_change1)) %>% 
  mutate(total_diff = rowSums(select(., c("first_diff", "second_diff")), na.rm = TRUE)) %>% 
  mutate(change = case_when(total_diff < 0 ~ "decrease",
         total_diff > 0 ~ "increase",
         total_diff == 0 ~ "no_change")
  )



changes_mtvernon <- mtvernon %>%
  select(neighborhood, address, total_diff, change) %>% 
  group_by(change) %>% 
  summarise(count = n(), avg_total_diff = mean(total_diff, na.rm = TRUE)) %>% 
  mutate(pct = round(count / sum(count), 2)) %>% 
  select(change, count, pct, avg_total_diff) %>% 
  mutate(neighborhood = "mtvernon")

#In Mount Vernon, 57% of the parking meters showed a decrease of an average of 43 cents. Just 22% of meters showed an increase of an average of 36 cents.

```

### Extract pdf table for Fells Point - use Tabula

--convert dollars to numbers --../data/FellsPointRateTables.csv

#Import neighborhood, create name, address and clean

```{r}
#rsw: added this to reset google sheets
googlesheets4::gs4_deauth()

fells <- read_sheet("https://docs.google.com/spreadsheets/d/1FOL2vBfA4ifBbbjjjQEVDCHklcI_OWvd3vghDiBqDWE/edit#gid=1640964516", "Fells") %>% 
  as.data.frame() %>% 
  clean_names() %>% 
  mutate(neighborhood = "fellspt") %>% 
  mutate(address = paste0(block, " ", street, ", ","Baltimore", ", ","Md")) %>% 
  mutate(rate1_day = as.numeric(first_rate_change_daytime_rate)) %>% 
   mutate(rate1_eve = as.numeric(first_rate_change_evening_rate)) %>% 
   mutate(rate2_day = as.numeric(second_rate_change_daytime_rate)) %>% 
   mutate(rate2_eve = as.numeric(second_rate_change_evening_rate))

#figure out why 800 S Broadway has $2.25 (Year 2022) 

#need to geocode the streets


glimpse(fells)
```

### Fells analysis

```{r}

fells_day <- fells %>% 
   mutate(first_day_diff= (rate1_day - initial_rate)) %>% 
  mutate(second_day_diff = (rate2_day - rate1_day)) %>% 
  mutate(total_diff_day = rowSums(select(., c("first_day_diff", "second_day_diff")), na.rm = TRUE)) %>% 
  mutate(change = case_when(total_diff_day < 0 ~ "decrease",
         total_diff_day > 0 ~ "increase",
         total_diff_day == 0 ~ "no_change"))

changes_fells_day <- fells_day %>%
group_by(change) %>%
  summarise(count_day = n(), price_diff_day = mean(total_diff_day, na.rm = TRUE)) %>%
  mutate(pct_day = round(count_day/sum(count_day), digits = 2)) %>% 
  mutate(avg_price_diff_day = round(price_diff_day, digits = 2)) %>% 
  select(change, count_day, pct_day, avg_price_diff_day) 


#I changed the summarize to be the mean of the total because before it was the mean of the first rate change.
  # group_by(change) %>%
 # summarise(count_day = n(), avg_first_day_diff = mean(first_day_diff, na.rm = TRUE)) %>%
#  mutate(pct_day = round(count_day/sum(count_day), digits = 2)) %>% 
#  select(change, count_day, pct_day, avg_first_day_diff) 

fells_eve <- fells %>% 
  mutate(first_eve_diff = (rate1_eve - initial_rate)) %>%
 mutate(second_eve_diff = (rate2_eve - rate1_eve)) %>% 
  mutate(total_diff_eve = rowSums(select(., c("first_eve_diff", "second_eve_diff")), na.rm = TRUE)) %>% 
  mutate(change = case_when(total_diff_eve < 0 ~ "decrease",
         total_diff_eve > 0 ~ "increase",
         total_diff_eve == 0 ~ "no_change")
  )

changes_fells_eve <- fells_eve %>%
   group_by(change) %>%
  summarise(count_eve = n(), price_diff_eve = mean(total_diff_eve, na.rm = TRUE)) %>%
  mutate(pct_eve = round(count_eve/sum(count_eve), digits = 2)) %>% 
  mutate(avg_price_diff_eve = round(price_diff_eve, digits = 2)) %>% 
  select(change, count_eve, pct_eve, avg_price_diff_eve) 
#I changed the summarize to be the mean of the total because before it was the mean of the first rate change

changes_fells <- changes_fells_day %>% 
  inner_join(changes_fells_eve, by=c("change")) %>% 
   mutate(neighborhood = "fellspt")




glimpse(changes_fells)


#Fix: i calculated avg_first_eve_diff1 as a separate column and then rounded it as a new column avg_first_eve_diff. there was some weirdness trying to round it back to the original. and I fixed the syntax on round to have , digits =2 follow it.


#the original that over-aggregated

#fells <- fells %>% 
# fedhill <- fedhill %>%
  # mutate(first_day_diff= (rate1_day - initial_rate)) %>% 
 # mutate(first_eve_diff = (rate1_eve - initial_rate)) %>%
 # mutate(second_day_diff = (rate2_day - rate1_day)) %>% 
 # mutate(second_eve_diff = (rate2_eve - rate1_eve)) %>% 
 # mutate(change = case_when(first_day_diff < 0 ~ "decrease",
       #  first_day_diff > 0 ~ "increase",
      #   first_day_diff == 0 ~ "no_change",
      #   first_eve_diff < 0 ~ "decrease",
     #    first_eve_diff > 0 ~ "increase",
      #   first_eve_diff == 0 ~ "no_change",
     #    second_day_diff < 0 ~ "decrease",
      #   second_day_diff > 0 ~ "increase",
   #      second_day_diff == 0 ~ "no_change",
    #     second_eve_diff < 0 ~ "decrease",
    #     second_eve_diff > 0 ~ "increase",
     #    second_eve_diff == 0 ~ "no_change"))


# changes_fells <- fells %>%
#   group_by(change) %>%
#   summarise(count = n(), avg_first_day_diff = mean(first_day_diff, na.rm = TRUE), avg_first_eve_diff1 = mean(first_eve_diff, na.rm = TRUE)) %>%
#   mutate(pct = round(count/sum(count), digits = 2)) %>% 
#   mutate(avg_first_eve_diff = round(avg_first_eve_diff1, digits = 2)) %>% 
#   select(change, count, pct, avg_first_day_diff, avg_first_eve_diff) %>%
#   mutate(neighborhood = "fellspt")



```

#Import Central Downtown

```{r}
cdowntown6_11 <- read_sheet("https://docs.google.com/spreadsheets/d/1FOL2vBfA4ifBbbjjjQEVDCHklcI_OWvd3vghDiBqDWE/edit#gid=1640964516", "Central Downtown6-11") %>% 
  as.data.frame() %>% 
  clean_names() %>% 
  mutate(address = paste0(block, " ", street, ", ","Baltimore", ", ","Md")) 

cdowntown1_5 <- read_sheet("https://docs.google.com/spreadsheets/d/1FOL2vBfA4ifBbbjjjQEVDCHklcI_OWvd3vghDiBqDWE/edit#gid=1640964516", "Central Downtown1-5") %>% 
  as.data.frame() %>% 
  clean_names() %>% 
  mutate(address = paste0(block, " ", street, ", ","Baltimore", ", ","Md")) %>% 
  mutate(initial_rate = as.numeric(as.character(starting_rate))) 

cdowntown <- cdowntown1_5 %>% 
  inner_join(cdowntown6_11, by="address") %>% 
  select(initial_rate, first_rate_change, second_rate_change, third_rate_change, fourth_rate_change, fifth_rate_change, sixth_rate_change, seventh_rate_change, eighth_rate_change, ninth_rate_change, tenth_rate_change, eleventh_rate_change, address) %>% 
  mutate(neighborhood = "cdowntown") 

cdowntown <- cdowntown %>% 
   mutate(first_rate1 = as.numeric(as.character(first_rate_change))) %>%
   mutate(second_rate1 = as.numeric(as.character(second_rate_change))) %>%
   mutate(third_rate1 = as.numeric(as.character(third_rate_change))) %>%
   mutate(fourth_rate1 = as.numeric(as.character(fourth_rate_change))) %>%
 mutate(fifth_rate1 = as.numeric(as.character(fifth_rate_change))) %>% 
  mutate(sixth_rate1 = as.numeric(as.character(sixth_rate_change))) %>%
   mutate(seventh_rate1 = as.numeric(as.character(seventh_rate_change))) %>%
   mutate(eigth_rate1 = as.numeric(as.character(eighth_rate_change))) %>%
   mutate(ninth_rate1 = as.numeric(as.character(ninth_rate_change))) %>%
   mutate(tenth_rate1 = as.numeric(as.character(tenth_rate_change))) %>%
 mutate(eleventh_rate1 = as.numeric(as.character(eleventh_rate_change)))
```

#Central Downtown analysis #RSW: NICE JOB

```{r}
cdowntown <- cdowntown %>%
  mutate(first_diff = (first_rate1 - initial_rate)) %>%
  mutate(second_diff = (second_rate1 - first_rate1)) %>%
   mutate(third_diff = (third_rate1 - second_rate1)) %>%
  mutate(fourth_diff = (fourth_rate1 - third_rate1)) %>%
    mutate(fifth_diff = (fifth_rate1 - fourth_rate1)) %>%
  mutate(sixth_diff = (sixth_rate1 - fifth_rate1)) %>%
  mutate(seventh_diff = (seventh_rate1 - sixth_rate1)) %>%
  mutate(eigth_diff = (eigth_rate1 - seventh_rate1)) %>%
  mutate(ninth_diff = (ninth_rate1 - eigth_rate1)) %>%
  mutate(tenth_diff = (tenth_rate1 - ninth_rate1)) %>%
  mutate(eleventh_diff = (eleventh_rate1 - tenth_rate1)) %>%
  mutate(total_diff = rowSums(select(., c("first_diff", "second_diff", "third_diff", "fourth_diff", "fifth_diff", "sixth_diff", "seventh_diff", "eigth_diff", "ninth_diff", "tenth_diff", "eleventh_diff")), na.rm = TRUE)) %>%
  mutate(change = case_when(total_diff < 0 ~ "decrease",
        total_diff > 0 ~ "increase", 
        total_diff == 0 ~ "no_change"))
  

changes_cdowntown <- cdowntown %>%
  select(neighborhood, address, total_diff, change) %>%
  group_by(change) %>%
  summarise(count = n(), avg_total_diff = mean(total_diff, na.rm = TRUE)) %>%
  mutate(pct = round(count / sum(count), 2)) %>%
  select(change, count, pct, avg_total_diff) %>%
  mutate(neighborhood = "cdowntown")

#In Central Downtown, 50% of meters saw an decreases at an average of $1.03. Just 44% saw a increases averaging $1.64.
```

#Import Fed Hill

```{r}

#rsw - I think I follow what you are saying. just verify it with the city

#Please see "Federal Hill" Column D on Google Sheet. They say $1.50/$2.00 but in "Fed Hill" I was able to logically assign the initial rate because if the rate adjustment one was recommended a decrease and the new price was $1.25, the initial had to have been $1.50.

fedhill <- read_sheet("https://docs.google.com/spreadsheets/d/1FOL2vBfA4ifBbbjjjQEVDCHklcI_OWvd3vghDiBqDWE/edit#gid=1640964516", "FedHill") %>%
  as.data.frame() %>%
  clean_names() %>%
  mutate(neighborhood = "fedhill") %>%
  mutate(address = paste0(block, " ", street, ", ","Baltimore", ", ","Md")) %>% 
  mutate(initial_day = as.numeric(initial_rate_daytime)) %>% 
    mutate(initial_eve = as.numeric(initial_rate_evening)) %>% 
  mutate(rate1_day = as.numeric(first_daytime_rate_change)) %>% 
   mutate(rate1_eve = as.numeric(first_evening_rate_change)) %>% 
   mutate(rate2_day = as.numeric(second_daytime_rate_change)) %>% 
   mutate(rate2_eve = as.numeric(second_evening_rate_change))


```

#Fed Hill analysis

```{r}

fedhill_day <- fedhill %>% 
   mutate(first_day_diff= (rate1_day - initial_day)) %>% 
  mutate(second_day_diff = (rate2_day - rate1_day)) %>% 
   mutate(total_diff_day = rowSums(select(., c("first_day_diff", "second_day_diff")), na.rm = TRUE)) %>% 
  mutate(change = case_when(total_diff_day < 0 ~ "decrease",
         total_diff_day > 0 ~ "increase",
         total_diff_day == 0 ~ "no_change")
  )

changes_fedhill_day <- fedhill_day %>%
  group_by(change) %>%
  summarise(count_day = n(), price_diff_day = mean(total_diff_day, na.rm = TRUE)) %>%
  mutate(pct_day = round(count_day/sum(count_day), digits = 2)) %>% 
  mutate(avg_price_diff_day = round(price_diff_day, digits = 2)) %>% 
  select(change, count_day, pct_day, avg_price_diff_day) 
#I changed the summarize to be the mean of the total because before it was the mean of the first rate change


fedhill_eve <- fedhill %>% 
  mutate(first_eve_diff = (rate1_eve - initial_eve)) %>%
 mutate(second_eve_diff = (rate2_eve - rate1_eve)) %>% 
   mutate(total_diff_eve = rowSums(select(., c("first_eve_diff", "second_eve_diff")), na.rm = TRUE)) %>% 
  mutate(change = case_when(total_diff_eve < 0 ~ "decrease",
         total_diff_eve > 0 ~ "increase",
         total_diff_eve == 0 ~ "no_change")
  )

changes_fedhill_eve <- fedhill_eve %>%
  group_by(change) %>%
  summarise(count_eve = n(), price_diff_eve = mean(total_diff_eve, na.rm = TRUE)) %>%
  mutate(pct_eve = round(count_eve/sum(count_eve), digits = 2)) %>% 
  mutate(avg_price_diff_eve = round(price_diff_eve, digits = 2)) %>% 
  select(change, count_eve, pct_eve, avg_price_diff_eve) 
#I changed the summarize to be the mean of the total because before it was the mean of the first rate change

changes_fedhill <- changes_fedhill_day %>% 
  inner_join(changes_fedhill_eve, by=c("change")) %>% 
  mutate(neighborhood = "fedhill")









#fedhill <- fedhill %>%
#  mutate(first_day_diff = (rate1_day - initial_rate)) %>%
 # mutate(first_eve_diff = (rate1_eve - initial_rate)) %>%
#  mutate(second_day_diff =(rate2_day - rate1_day)) %>%
#  mutate(second_eve_diff = (rate2_eve - rate1_eve)) %>%
#  mutate(total_diff = rowSums(select(., c("first_day_diff", "first_eve_diff", "second_day_diff", "second_eve_diff")), na.rm = TRUE,)) %>%
#  mutate(day_diff = rowSums(select(., c("first_day_diff","second_day_diff")), na.rm = TRUE,)) %>%
# mutate(evening_diff = rowSums(select(., c("first_eve_diff", "second_eve_diff")), na.rm = TRUE,)) %>%
    # mutate(change = case_when(first_day_diff < 0 ~ "decrease",
    #                         first_day_diff > 0 ~ "increase",
    #                         first_day_diff == 0 ~ "no_change",
    #                         first_eve_diff < 0 ~ "decrease",
    #                         first_eve_diff > 0 ~ "increase",
    #                         first_eve_diff == 0 ~ "no_change",
    #                         second_day_diff < 0 ~ "decrease",
    #                         second_day_diff > 0 ~ "increase",
    #                         second_day_diff == 0 ~ "no_change",
    #                         second_eve_diff < 0 ~ "decrease",
    #                         second_eve_diff > 0 ~ "increase",
    #                         second_eve_diff == 0 ~ "no_change")) 
#  mutate(change = case_when(day_diff < 0 ~ "decrease",
                     #       day_diff > 0 ~ "increase",
                       #     day_diff == 0 ~ "no_change",
                     #       evening_diff < 0 ~ "decrease",
                   #         evening_diff > 0 ~ "increase",
                 #           evening_diff == 0 ~ "no_change"))


#rsw fix - applied the same changes as in the fells point data: i calculated day_diff_avg1 as a separate column and then rounded it as a new column day_diff_avg. there was some weirdness trying to round it back to the original. and I fixed the syntax on round to have , digits =2 follow it.
#changes_fedhill <- fedhill %>%
#  group_by(change) %>%
#  summarise(count = n(), day_diff_avg1 = mean(first_day_diff, na.rm = TRUE), evening_diff_avg1 = mean(first_eve_diff, na.rm = TRUE)) %>%
#  mutate(pct = round(count/sum(count), digits = 2)) %>% 
 # mutate(evening_diff_avg = round(evening_diff_avg1, digits = 2)) %>% 
 # mutate(day_diff_avg = round(day_diff_avg1, digits = 2)) %>% 
#  select(change, count, pct, day_diff_avg, evening_diff_avg) %>%
#  mutate(neighborhood = "fedhill")

#I'm not sure whyt in summarise below, I get an error saying that object "day_diff" not found. I tried to unlist it as a numeric above to fix that, but it didn't work.
# 
# fedhill <- fedhill %>%   
#   mutate(day_diff = as.numeric(unlist(day_diff))) %>%
#   mutate(evening_diff = as.numeric(unlist(evening_diff)))
# 
# changes_fedhill <- fedhill %>%
#   select(neighborhood, address, first_day_diff, first_eve_diff, change) %>%
#   group_by(change) %>%
#   summarise(count = n(), avg_first_day_diff = mean(day_diff, na.rm = TRUE), avg_first_eve_diff = round(mean(evening_diff, na.rm = TRUE)),2) %>%
#   mutate(pct = round(count / sum(count)),2) %>%
#   select(change, count, pct, avg_first_day_diff, avg_first_eve_diff) %>%
#   mutate(neighborhood = "fedhill")

```

#rsw - go ahead and fix harbor east using the same strategy I applied
for fedhill and fells point

#Import Harbor East

```{r}
harboreast <- read_sheet("https://docs.google.com/spreadsheets/d/1FOL2vBfA4ifBbbjjjQEVDCHklcI_OWvd3vghDiBqDWE/edit#gid=1640964516", "Harbor East") %>%
as.data.frame() %>%
clean_names() %>%
mutate(neighborhood = "harboreast") %>%
mutate(address = paste0(block, " ", street, ", ","Baltimore", ", ","Md")) %>%
mutate(initial_rate = as.numeric(unlist(initial_rate))) 

as.data.frame(harboreast)

harboreast <- harboreast %>%
  mutate(rate1_day = as.numeric(first_rate_change_daytlme_rate)) %>%
  mutate(rate1_eve = as.numeric(first_rate_change_evening_rate)) %>%
  mutate(rate2_day = as.numeric(second_rate_change_daytlme_rate)) %>%
  mutate(rate2_eve = as.numeric(second_rate_change_evening_rate)) %>%
  mutate(rate3_day = as.numeric(third_rate_change_daytime_rate)) %>%
  mutate(rate3_eve = as.numeric(third_rate_change_even_ing_rate))
```

#Harbor East analysis

```{r}
harboreast_day <- harboreast %>% 
   mutate(first_day_diff= (rate1_day - initial_rate)) %>% 
  mutate(second_day_diff = (rate2_day - rate1_day)) %>% 
  mutate(third_day_diff =(rate3_day - rate2_day)) %>%
   mutate(total_diff_day = rowSums(select(., c("first_day_diff", "second_day_diff", "third_day_diff")), na.rm = TRUE)) %>% 
  mutate(change = case_when(total_diff_day < 0 ~ "decrease",
         total_diff_day > 0 ~ "increase",
         total_diff_day == 0 ~ "no_change")
  )

changes_harboreast_day <- harboreast_day %>%
  group_by(change) %>%
  summarise(count_day = n(), price_diff_day = mean(total_diff_day, na.rm = TRUE)) %>%
  mutate(pct_day = round(count_day/sum(count_day), digits = 2)) %>% 
  mutate(avg_price_diff_day = round(price_diff_day, digits = 2)) %>% 
  select(change, count_day, pct_day, avg_price_diff_day) 


harboreast_eve <- harboreast %>% 
  mutate(first_eve_diff = (rate1_eve - initial_rate)) %>%
 mutate(second_eve_diff = (rate2_eve - rate1_eve)) %>% 
  mutate(third_eve_diff = (rate3_eve - rate2_eve)) %>%
   mutate(total_diff_eve = rowSums(select(., c("first_eve_diff", "second_eve_diff", "third_eve_diff")), na.rm = TRUE)) %>% 
  mutate(change = case_when(total_diff_eve < 0 ~ "decrease",
         total_diff_eve > 0 ~ "increase",
         total_diff_eve == 0 ~ "no_change")
  )

changes_harboreast_eve <- harboreast_eve %>%
  group_by(change) %>%
  summarise(count_eve = n(), price_diff_eve = mean(total_diff_eve, na.rm = TRUE)) %>%
  mutate(pct_eve = round(count_eve/sum(count_eve), digits = 2)) %>% 
  mutate(avg_price_diff_eve = round(price_diff_eve, digits = 2)) %>% 
  select(change, count_eve, pct_eve, avg_price_diff_eve) 
#I changed the summarize to be the mean of the total because before it was the mean of the first rate change

changes_harboreast <- changes_harboreast_day %>% 
  inner_join(changes_harboreast_eve, by=c("change")) %>% 
  mutate(neighborhood = "harboreast")



#before splitting evening and daytime
#harboreast <- harboreast %>%
#  mutate(first_day_diff = (rate1_day - initial_rate)) %>%
 # mutate(first_eve_diff = (rate1_eve - rate1_eve)) %>%
 # mutate(second_day_diff =(rate2_day - rate1_day)) %>%
#  mutate(second_eve_diff = (rate2_eve - rate1_eve)) %>%
#  mutate(third_day_diff =(rate3_day - rate2_day)) %>%
#  mutate(third_eve_diff = (rate3_eve - rate2_eve)) %>%
#  mutate(day_diff = rowSums(select(., c("first_day_diff","second_day_diff", "third_day_diff")), na.rm = TRUE,)) %>%
#  mutate(evening_diff = rowSums(select(., c("first_eve_diff", "second_eve_diff", "third_eve_diff")), na.rm = TRUE,)) %>%
 # mutate(change = case_when(first_day_diff < 0 ~ "decrease",
                      #      first_day_diff > 0 ~ "increase",
                       #     first_day_diff == 0 ~ "no_change",
                       #     first_eve_diff < 0 ~ "decrease",
                        #    first_eve_diff > 0 ~ "increase",
                      #      first_eve_diff == 0 ~ "no_change",
                     #       second_day_diff < 0 ~ "decrease",
                      #      second_day_diff > 0 ~ "increase",
                   #         second_day_diff == 0 ~ "no_change",
                    #        second_eve_diff < 0 ~ "decrease",
                    #        second_eve_diff > 0 ~ "increase",
                  #          second_eve_diff == 0 ~ "no_change",
                   #         third_day_diff < 0 ~ "decrease",
                  #          third_day_diff > 0 ~ "increase",
                  #          third_day_diff == 0 ~ "no_change",
                  #          third_eve_diff < 0 ~ "decrease",
                   #         third_eve_diff > 0 ~ "increase",
                   #         third_eve_diff == 0 ~ "no_change"))

#changes_harboreast <- harboreast %>%
#  select(neighborhood, address, first_day_diff, first_eve_diff, change) %>%
#  group_by(change) %>%
#  summarise(count = n(), avg_day_diff = mean(day_diff, na.rm = TRUE), avg_eve_diff = round(mean(eve_diff, na.rm = TRUE)),2) %>%
#  mutate(pct = round(count / sum(count)),2) %>%
#  select(change, count, pct, avg_day_diff, avg_eve_diff) %>%
 # mutate(neighborhood = "harboreast")
```

### Note on the brute force conversion of the first_rate_change1:

Yes, both methods should achieve the same result. They both convert the
elements of the first_rate_change column to numeric values. The main
difference lies in the approach to handle potential size discrepancies
in the vectors within the column.

Your approach directly converts the column to character and then to
numeric. This assumes that the data in the column can be directly
converted to numeric without considering potential issues such as
missing or inconsistent values.

The alternative approach I provided handles potential issues with
varying vector lengths in the column. It pads shorter vectors with NA
values to ensure all vectors have the same length before converting them
to numeric. This approach is more robust if there are variations in the
length of the vectors within the column.

If you're confident that the data in your column is consistent and
doesn't have missing or inconsistent values, your approach of directly
converting to character and then to numeric is perfectly valid and
simpler.

Robot method

```{r}
 # Find the maximum length of elements in first_rate_change column
 max_length <- max(sapply(mtvernon$first_rate_change, length))

 #Pad shorter vectors with NA values to make them consistent
 
 mtvernon$first_rate_change <- lapply(mtvernon$first_rate_change, function(x) {
   length(x) <- max_length
   x
 })
```


#Select latest, and bind for day
```{r}
cdowntown_1 <- cdowntown %>% 
  select(address, neighborhood, eleventh_rate1) %>% 
  rename(latest_day = eleventh_rate1)

fedhill_1 <- fedhill %>% 
  select(address, neighborhood, rate2_day) %>% 
  rename(latest_day = rate2_day)


fells_1 <- fells %>% 
  select(address, neighborhood, rate2_day) %>% 
  rename(latest_day = rate2_day)
#**will need to be updated**

harboreast_1 <- harboreast %>% 
   select(address, neighborhood, fourth_rate_change_even_ing_rate) %>% 
  rename(latest_day = fourth_rate_change_even_ing_rate) %>% 
  mutate(latest_day = as.numeric(as.character(latest_day)))

mtvernon_1 <- mtvernon %>% 
   select(address, neighborhood, second_rate_change1) %>% 
  rename(latest_day = second_rate_change1)

parking_day <- rbind(mtvernon_1, harboreast_1, fells_1, fedhill_1, cdowntown_1)

```

#Geocode
```{r}

parking_day <- parking_day %>% 
   mutate(location = geocode(address))

write.csv(parking_day, "parking_day.csv")

```


#NIGHT
```{r}
cdowntown_2 <- cdowntown %>% 
  select(address, neighborhood, eleventh_rate1) %>% 
  rename(latest_day = eleventh_rate1)

fedhill_2 <- fedhill %>% 
  select(address, neighborhood, rate2_day) %>% 
  rename(latest_day = rate2_day)


fells_2 <- fells %>% 
  select(address, neighborhood, rate2_day) %>% 
  rename(latest_day = rate2_day)
#**Fells will need to be updated**

harboreast_2 <- harboreast %>% 
   select(address, neighborhood,fourth_rate_change_even_ing_rate) %>% 
  rename(latest_day = fourth_rate_change_even_ing_rate)

mtvernon_2 <- mtvernon %>% 
   select(address, neighborhood, second_rate_change1) %>% 
  rename(latest_day = second_rate_change1)

parking_night <- rbind(mtvernon_2, harboreast_2, fells_2, fedhill_2, cdowntown_2)

parking_night <- parking_night %>% 
   mutate(location = geocode(address))

write.csv(parking_night, "parking_night.csv")

```


#ALL CHANGES DAY
```{r}
cdowntown_3 <- cdowntown %>% 
  select(address, neighborhood, eleventh_rate1, change) %>% 
  rename(latest_day = eleventh_rate1)

fedhill_3 <- fedhill_day %>% 
  select(address, neighborhood, rate2_eve, change) %>% 
  rename(latest_day = rate2_eve)


fells_3 <- fells_day %>% 
  select(address, neighborhood, rate2_eve, change) %>% 
  rename(latest_day = rate2_eve)
#**Fells will need to be updated**

mtvernon_3 <- mtvernon %>% 
   select(address, neighborhood, second_rate_change1, change) %>% 
  rename(latest_day = second_rate_change1)

harboreast_3 <- harboreast_day %>% 
   select(address, neighborhood, fourth_rate_change_even_ing_rate, change) %>% 
  rename(latest_day = fourth_rate_change_even_ing_rate)

change_day <- rbind(mtvernon_3, fells_3, harboreast_3, fedhill_3, cdowntown_3) %>% 
  group_by(change) %>% 
   summarise(count_day = n())
```