---
title: "Baltimore Income Map"
author: "Rob Wells"
date: "7/6/2022"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

Median Income Maps By Census Tract in Baltimore Source: Kyle Walker's
tidycensus tutorial, Rachel Logan's leaflet map code for the Howard
Center

```{r message=FALSE, warning=FALSE}
#remotes::install_github("walkerke/tidycensus")
library(tidyverse)
library(tidycensus)
#a = get_decennial(geography = "state", variables = "P1_001N", year = 2020)
# head(a, 5)
#install.packages("formattable")
library(formattable)
library(htmlwidgets)
library(leaflet)
library(sf)
library(formattable)
```

Census API key.

```{r message=FALSE}
census_api_key("9cabe8a191a1f824755d4a1845f13cb08faa2c5f", install = TRUE)
```

There are two major functions implemented in tidycensus: get_decennial,
which grants access to the 1990, 2000, and 2010 decennial US Census APIs
get_acs, which grants access to the 5-year American Community Survey
APIs.

```{r}
#Calls variables for the 2020 decennial census
v20_dec <- load_variables(2020, "pl", cache = TRUE)
#Calls variables for the 2020 ACS census
v20_acs <- load_variables(2020, "acs5", cache = TRUE)
```

# Census Tract Median Income Using ACS

## Question: The 2010 and 2016 Maryland files each have 2812 rows, but the 2020 state file has 2950 rows. Why?

```{r}
#2020 Median Income By Census Tract ACS
#2016-2020
#B19001 COUNTS THE NUMBER OF HOUSEHOLDS
#B19013_001 is the household median income
md_income2020 <- get_acs(geography = "tract", 
              variables = c(number_households = "B19001_001", median_income = "B19013_001"), 
               state = "MD", 
               year = 2020) %>% 
  mutate(year=("2020"))

 
md_income2020
#B19013 defined: https://www.socialexplorer.com/data/ACS2010_5yr/metadata/?ds=ACS10_5yr&var=B19013001
```

```{r}
#2016 Median Income By Census Tract ACS
#2016-2012
md_income2016 <- get_acs(geography = "tract", 
              variables = c(number_households = "B19001_001", median_income = "B19013_001"), 
               state = "MD", 
               year = 2016) %>% 
  mutate(year=("2016"))
 
md_income2016
```

```{r}
#2010 Median Income By Census Tract 2010
#2006-2010 sample
md_income2010 <- get_acs(geography = "tract", 
              variables = c(number_households = "B19001_001", median_income = "B19013_001"), 
               state = "MD", 
               year = 2010) %>% 
  mutate(year=("2010"))
 
md_income2010
```

## Create Baltimore city-county, city files

```{r}
#Bind together 2010, 2016, 2020 files

md_income_all <- rbind(md_income2010, md_income2016, md_income2020)

md_income_all <- separate(data = md_income_all, col = NAME, into = c("Census_Tract", "County", "State"), sep = ",", extra = "merge", fill = "right")

#write.csv(md_income_all, "md_income_all.csv")

#Subset Baltimore city-county, city files
balt_income_all <- filter(md_income_all, grepl ("Baltimore", County))
#write.csv(balt_income_all, "balt_income_all.csv")

#Baltimore city median income, 2010, 2016, 2020

baltcity_income <- filter(balt_income_all, grepl ("Baltimore city", County)) %>% 
  filter(variable=="median_income")

```

```{r}
#reshape baltcity_income table

baltcity_income <- baltcity_income %>% 
  select(-moe) %>% 
  pivot_wider(names_from = "year", values_from = "estimate")

baltcity_income <- janitor::clean_names(baltcity_income)

baltcity_income <- baltcity_income %>% 
  mutate(Diff_2020_2010 = (x2020-x2010)) %>% 
  mutate(Diff_2020_2016 = (x2020-x2016))

# write.csv(baltcity_income, "baltcity_income.csv")
# write_rds(baltcity_income, "baltcity_income.rds")
```

# Map Median Income by Census Tract

```{r}
#subset the geometry
md_income2020_geo <- get_acs(geography = "tract", 
              variables = c(number_households = "B19001_001", median_income = "B19013_001"), 
               state = "MD", 
               year = 2020, geometry = TRUE) %>% 
  mutate(year=("2020")) %>% 
  select(GEOID, NAME)

#join geocoordinates with baltcity
baltcity_income <- md_income2020_geo  %>% 
 right_join(baltcity_income, by=c("GEOID"="geoid")) %>% 
  distinct()

#join Community Statistical Area names
#associates neighborhood names to Census Tracts using Community Statistical Areas, 2010 https://bniajfi.org/mapping-resources/
#Baltimore has 56 neighborhoods as measured by Community Statistical Areas
csa <- rio::import("https://raw.githubusercontent.com/profrobwells/Baltimore/main/CTA_CSA_2010.csv")

csa$GEOID10 <- as.character(csa$GEOID10)

baltcity_income <- baltcity_income  %>% 
 right_join(csa, by=c("GEOID"="GEOID10")) %>% 
  distinct()

#Rename Columns
baltcity_income <- baltcity_income %>% 
  rename(Neighborhood = CSA2010,
  Census = NAME10)

write_rds(baltcity_income, "baltcity_income.rds")
write.csv(baltcity_income, "baltcity_income.csv")
#Census Tracts 1801, 1802 have no data for 2020, but have data for 2010, 2016

```


```{r}
#Map by Median Income Difference
baltcity_income <- baltcity_income %>% 
  sf::st_transform('+proj=longlat +datum=WGS84')

pal <- colorNumeric(
  palette = "inferno",
  domain = baltcity_income$Diff_2020_2010
)

#https://leaflet-extras.github.io/leaflet-providers/preview/
leafMap <- leaflet() %>%
  addProviderTiles(providers$Esri.WorldStreetMap) %>%
  addPolygons(data = baltcity_income,
              color = ~pal(Diff_2020_2010),
              weight = 2.5,
              smoothFactor = 0.2,
              fillOpacity = 0.5,
              label = paste("Change is:",(scales::dollar(baltcity_income$Diff_2020_2010)), "for",baltcity_income$Neighborhood)) %>% 
  addLegend(
    position = "bottomright",
    pal = pal,
    values = baltcity_income$Diff_2020_2010,
    title = "Median Income Change<br>Per Census Tract <br> 2010 to 2020"
  )

#saveWidget(leafMap, "leafMap.html")
leafMap
```

```{r}
#Top / bottom Census tracts
#For Baltimore Data Day presentation
#Census tracts: generally 1,200 and 8,000 people, optimum size of 4,000
x <- baltcity_income %>% 
  as.data.frame() %>% 
  select(Neighborhood, x2020, x2010, Diff_2020_2010, Census) %>% 
  top_n(10, Diff_2020_2010) %>%
  arrange(desc(Diff_2020_2010))

x$x2010 <- formattable::currency(x$x2010, "$", format = "d")
x$x2020 <- formattable::currency(x$x2020, "$", format = "d")
x$Diff_2020_2010 <- formattable::currency(x$Diff_2020_2010, "$", format = "d")
```

```{r}
library(knitr)
library(kableExtra)
# install.packages("knitr")
# install.packages("kableExtra")
#install.packages("magick")
#library(magick)    
x %>% 
  kbl() %>% 
    kable_paper(full_width = F) %>%
  column_spec(1, bold = T, border_right = T) %>%
  column_spec(4, width = "10em", background = "yellow") %>%     save_kable("/Volumes/GoogleDrive/My Drive/Baltimore Class/DataOutput/Census.pdf", density =1000)
```



```{r}
#race by tract
#join with map - white /nonwhite for tooltip
#how white / nonwhite changed during the decade by tract

#2020: B02001_001 Estimate!!Total RACE
race2020 <- get_acs(geography = "tract", 
              variables = c(total = "B02001_001", white = "B02001_002", black = "B02001_003", hispanic = "B03001_003"), 
               state = "MD", 
               year = 2020) %>% 
  mutate(year=("2020"))

race2010 <- get_acs(geography = "tract", 
              variables = c(total = "B02001_001", white = "B02001_002", black = "B02001_003", hispanic = "B03001_003"), 
               state = "MD", 
               year = 2010) %>% 
  mutate(year=("2010"))
```

```{r}
md_race_all <- rbind(race2010, race2020)

md_race_all <- separate(data = md_race_all, col = NAME, into = c("Census_Tract", "County", "State"), sep = ",", extra = "merge", fill = "right")

#write.csv(md_income_all, "md_income_all.csv")

#Subset Baltimore city-county, city files
balt_race <- filter(md_race_all, grepl ("Baltimore", County))
#write.csv(balt_income_all, "balt_income_all.csv")

#Baltimore city median income, 2010, 2016, 2020

baltcity_race <- filter(balt_race, grepl ("Baltimore city", County)) 

```

```{r}
#reshape baltcity_income table

 baltcity_race2 <- baltcity_race %>% 
   dplyr::select(-moe, -State) %>% 
  pivot_wider(names_from = "year", values_from = "estimate")

baltcity_race2 <- janitor::clean_names(baltcity_race2)

 baltcity_race3 <- baltcity_race2 %>% 
  pivot_wider(names_from = "variable", values_from = c(x2010, x2020))

baltcity_race <- baltcity_race3 %>% 
  mutate(Pct_White_2020 = (x2020_white/x2020_total)) %>% 
  mutate(Pct_White_2010 = (x2010_white/x2010_total)) %>% 
  mutate(White_Pct_Diff_2020 = (Pct_White_2020-Pct_White_2010)) %>% 
  as.data.frame()

library(formattable)

baltcity_race$Pct_White_2010 <- formattable::percent(baltcity_race$Pct_White_2010)
baltcity_race$Pct_White_2020 <- formattable::percent(baltcity_race$Pct_White_2020)
baltcity_race$White_Pct_Diff_2020 <- formattable::percent(baltcity_race$White_Pct_Diff_2020)

write.csv(baltcity_race, "/Volumes/GoogleDrive/My Drive/Baltimore Class/baltcity_race.csv")

```


```{r}
#Revising data for the Fall class assignment
baltcity_income_clean <- baltcity_income %>% 
  as.data.frame() %>% 
  select(Neighborhood, x2010, x2016, x2020, Census, GEOID) %>% 
  arrange(Census)
write_csv(baltcity_income_clean, "baltcity_income_clean.csv")
```


```{r}
#2020-2010 Median Income By Maryland City ACS
#B19001 COUNTS THE NUMBER OF HOUSEHOLDS
#B19013_001 is the household median income
city_md_income2020 <- get_acs(geography = "place", 
              variables = c(number_households = "B19001_001", median_income = "B19013_001"), 
               state = "MD", 
               year = 2020) %>% 
  mutate(year=("2020"))

 
city_md_income2020
#B19013 defined: https://www.socialexplorer.com/data/ACS2010_5yr/metadata/?ds=ACS10_5yr&var=B19013001

city_md_income2020 <- separate(data = city_md_income2020, col = NAME, into = c("Place", "State"), sep = ",", extra = "merge", fill = "right")

#2010
city_md_income2010 <- get_acs(geography = "place", 
              variables = c(number_households = "B19001_001", median_income = "B19013_001"), 
               state = "MD", 
               year = 2010) %>% 
  mutate(year=("2010"))

city_md_income2010 <- separate(data = city_md_income2010, col = NAME, into = c("Place", "State"), sep = ",", extra = "merge", fill = "right")
#combine
city_md_income <- rbind(city_md_income2010, city_md_income2020)
#reshape, pivot
city_md_income <- city_md_income %>% 
  select(-moe) %>% 
  pivot_wider(names_from = "year", values_from = "estimate")

number <- city_md_income %>%
    filter(variable =="number_households")

median <- city_md_income %>%
    filter(variable =="median_income")

Xcity_md_income <- number %>% 
  inner_join(median, by=c("GEOID", "Place", "State")) %>%  
  janitor::clean_names() 

city_md_income <- Xcity_md_income %>%
    rename(households = variable_x, household_2010 = x2010_x, household_2020 = x2020_x, median_income = variable_y, median_inc_2010 = x2010_y, median_inc_2020 = x2020_y)

write.csv(city_md_income, "city_md_income.csv")
```



```{r}
#Analyze Md cities median income


y <- city_md_income %>% 
  mutate(Diff_Income = (median_inc_2020-median_inc_2010)) %>% 
  mutate(Pct_Income_Chg = (median_inc_2020-median_inc_2010)/median_inc_2010) %>% 
  mutate(Diff_Pop = (household_2020-household_2010)) %>% 
  mutate(Pct_Pop = (household_2020-household_2010)/household_2010) %>% 
  filter(household_2020 > 3512) %>% 
  select(place, Diff_Income, Pct_Income_Chg, median_inc_2010, median_inc_2020) %>% 
  arrange(desc(Pct_Income_Chg))

y$Diff_Income <- formattable::currency(y$Diff_Income, "$", format = "d")
y$median_inc_2010 <- formattable::currency(y$median_inc_2010, "$", format = "d")
y$median_inc_2020 <- formattable::currency(y$median_inc_2020, "$", format = "d")
y$Pct_Income_Chg <- formattable::percent(y$Pct_Income_Chg, 1)

md_city_summary <- y

#Clean names
md_city_summary$place <- gsub("CDP", "", md_city_summary$place)
md_city_summary$place <- gsub("city", "", md_city_summary$place)
md_city_summary$place <- gsub("town", "", md_city_summary$place)


write.csv(md_city_summary, "md_city_summary.csv")

```

```{r}
#Chart of cities
ggplot(md_city_summary,aes(x = reorder(place, -Pct_Income_Chg), y = Pct_Income_Chg,
             fill = Pct_Income_Chg)) +
  geom_col(position = "dodge") + 
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=3))+
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Gain in Median Household Income in Maryland",
       subtitle = "2010 -> 2020 Census for Largest Cities, Towns",
       caption = "Graphic by Rob Wells, 7-19-2022",
       y="Pct Change 2010 vs 2020",
       x="Place")

ggsave("md_city_summary.png",device = "png",width=9,height=6, dpi=800)


```

```{r}
#Chart of cities, shortened

md_city_summary %>% 
  top_n(25, Pct_Income_Chg) %>% 
ggplot(aes(x = reorder(place, -Pct_Income_Chg), y = Pct_Income_Chg,
             fill = Pct_Income_Chg)) +
  geom_col(position = "dodge") + 
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=10))+
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Gain in Median Household Income in Maryland",
       subtitle = "2010 -> 2020 Census for Largest Cities, Towns",
       caption = "Graphic by Rob Wells, 7-19-2022",
       y="Pct Change 2010 vs 2020",
       x="Place")

ggsave("top_md_city_summary.png",device = "png",width=9,height=6, dpi=800)


```


###stopped July 19

###stopped July 19

###stopped July 19

next: link balt city race to the map and add tooltip

```{r}
#Overlay with HOLC redlining maps
#https://s3.amazonaws.com/holc/tiles/MD/Baltimore/1937/holc-scan.jpg
#install.packages("rjson")
library("rjson")
holc <- fromJSON(file="/Volumes/GoogleDrive/My Drive/Baltimore Class/Tracts_2020_HOLC/Tracts_2020_HOLC.geojson")

#install.packages("rgdal")
library(rgdal)
library(raster)
holc1 <- shapefile("/Volumes/GoogleDrive/My Drive/Baltimore Class/Tracts_2020_HOLC/Tracts_2020_HOLC.shp")

holc2 <- holc1@data %>% 
  filter(MAX_city=="Baltimore")

```

```{r}
#Notes on embedding and formatting Leaflet

#https://profrobwells.github.io/Baltimore/leafMap.html


#<iframe seamless src="/static/leaflet/leafmap/index.html" width="100%" height="500"></iframe>
#https://waterdata.usgs.gov/blog/leaflet/
#https://github.com/profrobwells/Test/blob/master/leafMap.html
#https://leaflet-extras.github.io/leaflet-providers/preview/


#notes on old code
#label = ~Diff_2020_2010) %>%
              #label = paste("Change is:  $",baltcity_income$Diff_2020_2010, ",",baltcity_income$census_tract)) %>% 

#next up

#Put population and %white in popup
#Get more details on how to format popups

```
