---
title: "Internet Access and Income in Baltimore"
name: Victoria Stavish
date: Oct 20, 2022
format: html
editor: visual
---

## Internet Access Improving Even in Baltimore's Struggling Neighborhoods

-   Internet access is improving in some Baltimore City neighborhoods, despite a decreasing median income.

-   While there is a high correlation between internet access and income, we found that some neighborhoods in Baltimore City are seeing gains in internet access even as median income decreases.

Even though some Baltimore City neighborhood saw median income fall over the last decade, internet access continues to improve among all Baltimore City neighborhoods. 

For example, the ten Baltimore neighborhoods with the no gains or the smallest gains in median income between 2016 and 2020, all but one saw significant decreases in the percentage of homes without internet access. 

Southwest Baltimore saw a 38% decrease in median income, the number of households with internet subscriptions rose 9% between 2017 and 2020. Upton/Druid Heights also saw their median income decrease by 17% between 2016 and 2020, but the percent of homes without internet also rose by nearly 17% between 2017 and 2020. 

We still have questions about these trends. 

While income clearly has an affect, why does a decrease in median income not necessarily mean a decrease in internet access? 
Are we seeing a shift towards low income correlating less and less with internet access?

Some places saw an increase in median income and a rise in the number of households without internet access. However, these places are few and the changes in those numbers are drastic. This makes it less noteworthy than neighborhoods in which internet access increased while median income decreased.

The general theme of internet access and income having a strong correlation is still true. In 2020, of the 10 neighborhoods in Baltimore City with the best internet access, the median income was about \$100,000, while in the neighborhoods in Baltimore City with the worst internet access, the median income was about \$34,000.

--Victoria Stavish

## Analysis

```{r include=FALSE}
# load libraries 

library(tidyverse)
library(janitor)
library(ggplot2)
library(dplyr)
library(ggthemes)
library(DT)

```

**Internet Access and Income in Baltimore:**

The goal of this notebook is to analyze the correlation between average incomes and internet access in Baltimore City neighborhoods. To do this I am using a dataset that includes the median income in Baltimore City neighborhoods based on 2020 Census data. I am also using an open source dataset from the City of Baltimore that details how many homes do not have internet subscriptions in each Baltimore City neighborhood.

**Questions:**

-   Do high income areas have high internet access?

-   Do low income areas have poor internet access?

-   Does internet access improve/get worse for places whose median income increased/decreased 2016-2020?

**Findings **

While neighborhoods with good internet access usually have higher median incomes and neighborhoods with low median incomes usually have poor internet access, there does not seem to be a significant association between decreasing median incomes and lower internet access overtime. Most neighborhoods increase internet access even when they saw negative percent change in median income.

** Caveats in the data**

In both datasets, some neighborhoods are repeated multiple times with different numbers in the same year, so it is unclear which number is the most accurate representation of the neighborhood. The internet subscription dataset also does not detail how they define an internet subscription(ie. broadband or wireless).

```{r include=FALSE}
# load income data

balt_income <- read_csv("baltcity_income_clean.csv")%>%
clean_names()

## Data dictionary
# This dataset has 6 columns. The first column is the names of Baltimore City neighborhoods (chr), the following three columns show the median income in each neighborhood in 2010 (dbl), 2016 (dbl) and 2020 (dbl). The fourth column shows the census id for each neighborhood (dbl) and the fifth column shows the geography id of each neighborhood (dbl). 

# load internet data

balt_internet <- read_csv("balt_internet.csv") %>%
  clean_names()

## Data dictionary
# This dataset has 9 columns. The first column gives an object id (dbl) to each row. The second column gives the Baltimore City neighborhood name as of the 2010 census data (chr). The following three columns give the percentage of households in each neighborhood that do not have an internet subscription for the years 2017 (dbl), 2018 (dbl) and 2019 (dbl). The sixth column gives the Baltimore City neighborhood name as of the 2020 census data (chr). The seventh column gives the percentage of households in each neighborhood that do not have an internet subscription for the year 2020 (dbl). The eighth and ninth columns give the neighborhood shape (dbl) and length (dbl), neither of which are relevant for our analyses. 

```

**Do low income neighborhoods have low percentage of internet access?**

**The ten neighborhoods with highest percentage of homes with no internet access, what is the median income?**

```{r}

# 10 neighborhoods with lowest internet access

balt_internet_b10 <- balt_internet %>%
  as.data.frame() %>%
  clean_names() %>%
 dplyr::: select(csa2020, nohhint20) %>%
  arrange(desc(nohhint20)) %>%
  slice_max(nohhint20, n=10)

# median percentage of homes without internet access among the 10 neighborhoods with the lowest internet access = 31.66%
summary(balt_internet_b10)
```

```{r}
# New DF combining the income and internet by the csa2010 names. The csa2020 names don't match. This produces 199 matches. It leaves out 	
# Unassigned--Jail, which is junk anyway

#Turned the Internet access into a positive value 

internet_income_all <- balt_income %>% 
  inner_join(balt_internet, by=c("neighborhood"="csa2010")) %>% 
  select(neighborhood, x2010, x2016, x2020, nohhint17, nohhint20) %>% 
  mutate(income_diff = x2020-x2016) %>% 
  mutate(income_pct_chg = formattable::percent(x2020-x2016)/x2016) %>% 
  mutate(internet2017 =abs(round(nohhint17-100))) %>% 
  mutate(internet2020 =abs(round(nohhint20-100))) %>% 
  mutate(internet_diff = internet2020-internet2017) %>% 
  mutate(internet_pct_chg =  formattable::percent(internet2020-internet2017)/internet2017)

x <- internet_income_all %>% 
  mutate(income_chg = round(income_pct_chg, 2)) %>% 
  mutate(internet_chg = round(internet_pct_chg, 2))  %>% 
  select(neighborhood, internet2020, internet_chg, x2016, x2020, income_chg)

datatable(x, options = list(
  autoWidth = TRUE,
  columnDefs = list(list(width = '1px', targets = c(0:6)))
))

#datatable formatting
#https://rstudio.github.io/DT/options.html
```

```{r}

internet_income2 <- balt_income %>% 
  inner_join(balt_internet, by=c("neighborhood"="csa2010")) %>% 
  select(neighborhood, x2010, x2020, nohhint17, nohhint20) %>% 
  group_by(neighborhood) %>% 
  summarize(internet2020 =abs(round(mean(nohhint20-100))))

#How we figured out what wasn't joined:
# internet2_income <- balt_income %>% 
#   anti_join(balt_internet, by=c("neighborhood"="csa2010"))

#Here we average the neighborhoods so places with multiple entries will have a single entry and a single median income figure.

internet_income <- balt_income %>% 
  inner_join(balt_internet, by=c("neighborhood"="csa2010")) %>% 
  select(neighborhood, x2010, x2020, nohhint17, nohhint20) %>% 
  group_by(neighborhood) %>% 
  summarize(income2020 = round(mean(x2020, na.rm=TRUE, 1))) %>% 
  inner_join(internet_income2, by=c("neighborhood")) 

datatable(internet_income, options = list(
  autoWidth = TRUE,
  columnDefs = list(list(width = '1px', targets = c(0:3)))
))

```

**In neighborhoods with lowest percentage of internet access, what is the median salary?**

```{r}
summary(internet_income$internet2020)
  #  Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  # 3.432   9.941  16.414  18.299  25.492  38.596 

b10_income <- internet_income %>% 
  filter(internet2020 > 25.492) %>% 
  arrange(desc(internet2020))

b10_income %>% 
  summarize(mean(income2020))

# median income in neighborhoods with lowest percentage of internet access (the lowest quartile for internet access, 26% and above) = $36,736.62				

```
**Do high income neighborhoods have high percentage of internet access?**

The ten neighborhoods with highest percentage of homes with internet access, what is the median income?

```{r}

# 14 neighborhoods with high internet access, or those in the upper quartile of access (less than 10% of the households lack access)

t10_balt_internet <- internet_income %>% 
  filter(internet2020 < 9.941) %>% 
  arrange(internet2020)

t10_balt_internet %>% 
  summarize(mean(income2020))

# median income in neighborhoods with highest percentage of internet access = $88,127.8

```

**Does internet access improve/get worse for places whose median income increased/decreased 2016-2020?**

```{r}
# top 10 neighborhoods with greatest growth in median income 2016-2020

internet_income_all <- internet_income_all %>% 
  mutate(income_pct_chg = (x2020 - x2016)/x2016*100) %>%
  mutate(internet_pct_chg = (nohhint20-nohhint17/nohhint17*100)) %>% 
  mutate(internet_chg = (nohhint20-nohhint17))


# poschange_income <- balt_income %>%
#   select(neighborhood, x2016, x2020) %>%
#   mutate(pct_change = ((x2020 - x2016)/x2016*100)) %>%
#   arrange(desc(pct_change)) %>%
#   slice_max(pct_change, n=10)

# Those 10 neighborhoods, how did their internet access change between 2017-2020

poschange_income <- internet_income_all %>% 
  slice_max(income_pct_chg, n=10)
#The larger the negative percentage change, the better the improvement in Internet access. For example, the -89.11368 for 	Greater Charles Village/Barclay meant it went from 81% to 91% of households with Internet access in four years

```

**A general look at changes in internet access compared to changes in median income in Baltimore City neighborhoods 2016-2020.**

```{r}
income_change <- balt_income %>%
  dplyr:::select(neighborhood, x2016, x2020) %>%
  mutate(income_pct_change = ((x2020 - x2016)/x2016*100)) 


# mutate the internet data to see percent change over time

internet_change <- balt_internet %>%
  clean_names() %>%
  dplyr:::select(csa2020, nohhint17, nohhint20) %>%
  mutate(internet_pct_change = ((nohhint20 - nohhint17)/nohhint17*100)) 
  
# join the change in internet dataframe with the change in income dataframe
  
change_join <- internet_change %>%
  inner_join(income_change, c("csa2020" = "neighborhood"))


```

**Does internet access improve/get worse for places whose median income increased/decreased 2016-2020?**

```{r}

# 10 neighborhoods with greatest decrease in median income 2016-2020

negchange_income <- internet_income_all %>% 
  slice_min(income_pct_chg, n=10)
#All of these neighborhoods that saw a drop in income still saw improvement in internet access, except for one part of Sandtown-Winchester/Harlem Park

```

**Scatterplot of internet access and income in neighborhoods in 2020**

```{r}
# Join the dataframes for Baltimore City income and Baltimore City internet

income_internet_join <- balt_income %>%
  clean_names() %>%
  group_by(neighborhood, x2020) %>%
  inner_join(balt_internet, c("neighborhood" = "csa2020")) %>%
  select(neighborhood, x2020, nohhint20)

# Create a scatterplot to show the relationship between income and internet access in Baltimore City neighborhoods

#In 2020 Baltimore City neighborhoods with higher median incomes generally had more home with internet subscriptions than neighborhoods with low median incomes.

ggplot(income_internet_join, aes(y=nohhint20, x=x2020, fill=nohhint20)) +
geom_point() +
theme(legend.position = "none") +

  labs(
    title="Internet access, income in Baltimore",
    subtitle = "Each point represents a neighborhood in Baltimore City,",
    x = "Median income",
    y = "Pct homes w/o internet access",
    caption = "Source: The U.S. Census Bureau  By: Victoria Stavish")

```
