---
title: "UMD Fees"
date: 9-08-2025
---
Your Name - Rob Wells
Follow along with the demo to learn how to calculate percentage change using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.


### Task 1: Load the tidyverse, janitor
Janitor performs important data cleaning operations

```{r}
library(tidyverse)
install.packages('janitor')
library(janitor)
```




### Task 2: Load the UMD Fees dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/umd_fees.csv


```{r}
umd_fees <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/umd_fees.csv") |> 
   clean_names()
```

### Task 3: Calculate the percentage change in each category between Fall 2024 and Fall 2021 using `mutate()`

```{r}
umd_fees_pct <- umd_fees |> mutate(pct_change = (fall_2024- fall_2021)/fall_2021 * 100)
```


### Task 4: Calculate the mean and median fees for Fall 2024 using `mutate()`

```{r}

umd_fees |> 
    summarize(mean_2021_fees = mean(fall_2021),
            median_2024_fees = median(fall_2024,
            mean_2024_fees = mean(fall_2021))
)


```

### Task 5: Add to the previous the max and minimum values for 2024 in addition to the mean and median fees for Fall 2024 using `mutate()`

```{r}

umd_fees |> 
    summarize(mean_2024_fees = mean(fall_2024),
            median_2024_fees = median(fall_2024),
            max_2024_fees = max(fall_2024),
            min_2024_fees = min(fall_2024)
)


```

### Mutate exercise

Mutate is a powerful command that allows you to create a new column and fill it with the results of another calculation. We will be using it a lot.

In this demonstration, we calculate the total number of incidents per year for each category using `mutate()`, calling the new column `total_fees`. 

Then calculate the average number of incidents per month for each category of crime, calling the new column `average_fees`. Assign the result to a new dataframe called `umd_fee_averages`.


```{r}

umd_fees_pct <- umd_fees |> 
  mutate(total_fees1 = fall_2024 + fall_2023 + fall_2022 + fall_2021) |> 
  mutate(total_fees = rowSums(across(2:5))) |> 
  mutate(average_fees = total_fees/4)

```

Note that lines 73 and 74 do the same thing. 
Line 74 is preferred because there's less typing! rowSums(across) tells R to add across the rows. We specify rows 2 through 5.


PRO TIP! Remember this mutate command for the homework.
HINT HINT HINT

### Task 6: Describe our results

Write a few sentences about what you found in the data. What's the story here?





When you are done, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
