---
title: "Baltimore Foot Traffic"
author: "Rob Wells"
date: "2025-01-22"
output: html_document
---
```{r}
library(tidyverse)
#remotes::install_github("walkerke/tidycensus")
library(tidyverse)
library(tidycensus)
#a = get_decennial(geography = "state", variables = "P1_001N", year = 2020)
# head(a, 5)
#install.packages("formattable")
library(formattable)
library(htmlwidgets)
#install.packages("leaflet")
library(leaflet)
library(sf)
library(formattable)
library(janitor)
#install.packages("leaflet.extras")
library(leaflet.extras)
library(googlesheets4)
```

It contains monthly foot traffic information for 20,000 stores in Baltimore City, Baltimore County, and Anne Arundel County from Jan to July 2024 (NAICS=44,45,71,72, 81). 
For each store, we have the aggregate foot traffic for the month, the same-month foot traffic last year, and the rate of change. You can also find information about the number of unique visitors and their proximity—whether they are from nearby or outside the county.

#Import data
```{r}
#foot <- rio::import("../data/baltimore_foot_traffic.dta") - old
foot1 <- rio::import("../data/baltimore_foot_traffic_feb21_2025.dta")
#write.csv(foot1, "baltimore_foot_traffic.csv")
```

```{r}
glimpse(foot)
```

Dundalk zip code 21222
Edgemere 21219, 21052
Brandon Shores 21226

# Dundalk / Arundel
```{r}
#1671 results
focus <- foot1 |> 
  filter(postal_code == c("21222", "21219","21052", "21226", "21225", "21060", "21061", "21090"))
focus |> 
  count(city) |> 
  arrange(desc(n))
```

#Biggest losers

```{r}
losers <- focus |> 
  select(location_name, street_address, city, region, postal_code, chg_nvisits_yoy, start, placekey) |> 
  group_by(placekey) |> 
  slice_max(start, n=1) |> #select the most recent data if multiple entries
  ungroup() |> 
  arrange(chg_nvisits_yoy) |> 
  slice_min(chg_nvisits_yoy, n = 200) |>
  mutate(address = paste0(street_address, " ", city, " ", region, " ", postal_code)) |> 
  mutate(visit_drop = formattable::percent(chg_nvisits_yoy))

#write.csv(losers, "dundalk_arundel_losers.csv")


#fact check

# losers |> 
# count(location_name) |> 
# arrange(desc(n))

```




#Geocode Dundalk2
```{r}
library(tidyr)
library(ggmap)
register_google(key = "")


losers <- losers %>% 
  mutate(geo = geocode(address)) 

losers <- losers %>%
  tidyr::unnest(geo)


losers <- st_as_sf(losers, coords = c("lon", "lat"), crs = 4326)

# write.csv(losers, "dundalk_arundel_losers.csv")
```


#Map 
```{r}
leafMap <- leaflet() %>%
  addProviderTiles(providers$Esri.WorldStreetMap) %>%
  setView(lng = -76.52, lat = 39.27, zoom = 11.5) %>%
  addCircleMarkers(
    data = losers,
    radius = 8,
    color = "black",
    weight = 1,
    fillColor = ~pal(chg_nvisits_yoy),
    fillOpacity = 0.7,
    popup = ~paste("<b>", location_name, "</b><br>",
                  "Change in foot traffic: ", visit_drop, "<br>",
                  "For the time period", start, "<br>",
                  "Address: ", address)
  ) |> 
addControl(
    html = "<div style='background-color: white; padding: 0.0001px; border-radius: 0.0001px; text-align: center;'><h4 style='margin-bottom: 2px;'>Yearly Business Decline, Dundalk & Anne Arundel Co., since Key Bridge Collapse</h4><h5 style='margin-top: 0px;'>Click dot for business details</h5></div>",
    position = "topright"
  )
leafMap
saveWidget(leafMap, "../output/foot_traffic_decline_map_feb_2025.html")
```


#Earlier notes


```{r}
arundel <- foot |> 
  filter(postal_code == c("21225", "21226","21060", "21061", "21090", "21122"))

arundel |> 
  count(city) |> 
  arrange(desc(n))


```


```{r}
arundel_loser <- arundel |> 
  select(location_name, street_address, city, region, postal_code, dln_visits_yoy, start) |> 
  arrange(dln_visits_yoy) |> 
  slice_min(dln_visits_yoy, n = 100) |>
  mutate(foot_traffic_yoy = abs(dln_visits_yoy)) |> 
  mutate(address = paste0(street_address, " ", city, " ", region, " ", postal_code)) 
  


write.csv(arundel_loser, "arundel_loser.csv")
  


```


# Table
```{r}
library(kableExtra)
table_output <- losers %>%
  kbl(caption = "Baltimore Businesses Losing Customers") %>%
  kable_classic(full_width = F, html_font = "Cambria") %>%
  column_spec(1, bold = T, border_right = T) %>%
  column_spec(2, bold = F, border_right = T) |> 
  column_spec(3, width = "5em", background = "yellow") |> 
column_spec(4, bold = F, border_right = T) 

cat("<style>
    .caption {
      font-size: 30px !important;
    }
    </style>")
table_output


```
#Geocode Foot Traffic

Build lookup table: 16,475 entries
Foot has multiple entries for the same address
A lookup table will link back to the main database while saving the expense of geocoding 143,000 entries
```{r}
foot_sum <- foot |> 
  mutate(fulladdress = paste(street_address, city, region, postal_code, sep = ", ")) |> 
  distinct(fulladdress, .keep_all = TRUE)


```

### Geocode lookup table
```{r}
library(tidyr)
library(ggmap)
register_google(key = "")


foot_sum <- foot_sum %>% 
  mutate(geo = geocode(fulladdress)) 
write.csv(foot_sum, "foot_traffic_geocode_lookup.csv")

foot_sum <- foot_sum %>% 
  unnest(geo)

#Join with main data

foot_sum_short <- foot_sum |> 
  select(fulladdress, lon, lat)

foot1 <- foot |> 
  mutate(fulladdress = paste(street_address, city, region, postal_code, sep = ", ")) |>
  left_join(foot_sum_short, by=c("fulladdress"))

```

fact check
```{r}
sum(is.na(foot1$lat))

### total the places without lat lon
x <- foot1 |> 
  group_by(city) |> 
  count(is.na(lat)) |> 
  arrange(desc(n)) |> 
  rename(valid = 'is.na(lat)')

foot1 |> 
  group_by(city) |> 
  count() |> 
  arrange(desc(n))

```
#failed spatial join with business data
```{r}
bol <- read.csv("../data/baltimore_bol.csv") |> 
  janitor::clean_names() |> 
  select(company_name,       address_1,          address_2,         
city,               state,              zip,                zip4,              
latitude,           longitude,          phone,              fax,               
website,            email,              employees,          revenue,           
contact_formal_name,first_name,        
middle,             last_name,          suffix,             title,             
     last_published_date, classification, ranking_criteria, list_title, company_id, business_info) 

```

# spatial join
```{r}
foot_traffic_biz <- foot1 |>
  inner_join(bol, by = c("lat" = "latitude", "lon" = "longitude"))

```

# Unemployment Data

```{r}
unem_feb <- rio::import("/Users/robwells/Downloads/Unemployment_Feb2024_March2024.xlsx", sheet="Feb2024" ) |> 
  janitor::clean_names() |> 
  rename(feb_count = count) |> 
   filter(state_code_value=="MD")

unem_march <- rio::import("/Users/robwells/Downloads/Unemployment_Feb2024_March2024.xlsx", sheet="March 2024") |> 
  janitor::clean_names() |> 
  rename(march_count = count) |> 
   filter(state_code_value=="MD")

combined_unem <- unem_feb |> 
  inner_join(unem_march, by=c("zip", "year", "state_code_value")) |> 
  mutate(diff_march_feb = (march_count-feb_count))

write.csv(combined_unem, "combined_unem_feb_march_2024.csv")

```

```{r}

unem_feb |> 
  count(state_code_value) |> 
  arrange(desc(n)) |> 
  filter(state_code_value=="MD")


```

```{r}
unem_march |> 
  count(state_code_value) |> 
  arrange(desc(n))
```

```{r}

combined_unem |> 
  count(state_code_value) |> 
  arrange(desc(n))
```

