---
title: "Smith School - Baltimore Data Income Map"
author: "Rob Wells"
date: "1/28/2024"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r message=FALSE, warning=FALSE}
#remotes::install_github("walkerke/tidycensus")
library(tidyverse)
library(tidycensus)
#a = get_decennial(geography = "state", variables = "P1_001N", year = 2020)
# head(a, 5)
#install.packages("formattable")
library(formattable)
library(htmlwidgets)
#install.packages("leaflet")
library(leaflet)
library(sf)
library(formattable)
library(janitor)
#install.packages("leaflet.extras")
library(leaflet.extras)
library(googlesheets4)

```

Census API key.

```{r message=FALSE}
census_api_key("9cabe8a191a1f824755d4a1845f13cb08faa2c5f", install = TRUE, overwrite = TRUE)
```

# --------------------------------------
# Maryland Data Import:  Smith School 
# --------------------------------------

```{r}
md <- rio::import("../data/maryland_tracts.xls") %>% 
  as.data.frame()

md <- md %>% 
  clean_names() %>% 
  mutate(census = as.character(tract)) %>% 
  mutate(county = as.numeric(strtrim(tract, 5))) %>% 
  mutate(geoid = as.numeric(strtrim(tract, 7))) %>% 
  mutate(geoid1 = as.character((geoid))) %>% 
  mutate(zip1 = as.character((zip)))
```


```{r}
mdzip <- rio::import("../data/ZIP_TRACT_122023.xlsx")



md_places <- md %>% 
  inner_join(mdzip, by=c("zip1"="ZIP")) %>% 
  distinct(tract, .keep_all = TRUE) %>% 
  mutate(pct_anymembershp1 = formattable::percent(pct_anymembershp_zip)) |> 
  janitor::clean_names() 


#write.csv(md_places, "../data/smith_md_places_april1.csv")
```

# Simply Smith Data for B'more class
```{r}
md_places_brief <- md_places |> 
   select(usps_zip_pref_city,  income_per_capita, census_response_rate2020, census_response_rate2010, nbanks_zip, pct_anymembershp_zip, zip, tot_pop, age_above65, median_age, white, black, asian, geoid1, tract) |> 
  rename(city = usps_zip_pref_city)

write.csv(md_places_brief, "smith_school_brief_version_1_27-2025.csv")

md_places_full <- md_places |> 
     select(usps_zip_pref_city,  income_per_capita, census_response_rate2020, census_response_rate2010, nbanks_zip, pct_anymembershp_zip, zip, tot_pop, age_above65, median_age, white, black, asian, edu_highschool, edu_bachelor,edu_graduate, income_0to10k, income_10to50k, income_50to100k, income_100to200k, income_200k, income_median, income_per_capita, occupation_management, occupation_sales, occupation_naturalresource, occupation_production, res_ratio, bus_ratio, oth_ratio,tot_ratio,pct_anymembershp1,geoid1, tract, census) |> 
  rename(city = usps_zip_pref_city)

write.csv(md_places_full, "smith_school_w_places_full_version_1_27-2025.csv")

```








# Census Tract Median Income Using ACS

#Subset geocoordinates
```{r}
#subset the geometry

#2020 Median Income By Census Tract ACS
#2016-2020
#B19001 COUNTS THE NUMBER OF HOUSEHOLDS
#B19013_001 is the household median income
md_income2020_geo <- get_acs(geography = "tract", 
              variables = c(number_households = "B19001_001", median_income = "B19013_001"), 
               state = "MD", 
               year = 2020, geometry = TRUE) %>% 
  mutate(year=("2020")) %>% 
  select(GEOID, NAME)

maryland_places_data <- get_acs(geography = "place",
                                 variables = c(number_households = "B19001_001", median_income = "B19013_001"),
                                 state = "MD",
                                 year = 2020)


# Retrieve ACS data for places in Maryland
places_data <- get_acs(geography = "place",
                        variables = c(number_households = "B19001_001", median_income = "B19013_001"),
                        state = "MD",
                        year = 2020)

# Retrieve ACS data for tracts in Maryland
# Ensure caching of shapefiles
options(tigris_use_cache = TRUE)
tracts_data <- get_acs(geography = "tract",
                        variables = c(number_households = "B19001_001", median_income = "B19013_001"),
                        state = "MD",
                        year = 2020,
                        geometry = TRUE)

# Merge the place and tract data
maryland_data <- bind_rows(places_data, tracts_data)

#B19013 defined: https://www.socialexplorer.com/data/ACS2010_5yr/metadata/?ds=ACS10_5yr&var=B19013001

md_named <- md_named %>% 
  mutate(GEOID = as.character(tract))

#join geocoordinates 
md2 <- md_income2020_geo  %>% 
 right_join(md_named, by=c("GEOID")) %>% 
 distinct()

md2 <- md2 %>%
   sf::st_transform('+proj=longlat +datum=WGS84')

```





```{r}
pal <- colorNumeric(
  palette = "inferno",
  domain = md2$pct_anymembershp_zip
)

#https://leaflet-extras.github.io/leaflet-providers/preview/
leafMap <- leaflet() %>%
  addProviderTiles(providers$Esri.WorldStreetMap) %>%
  addPolygons(data = md2,
              color = ~pal(md2$pct_anymembershp_zip),
              weight = 2.5,
              smoothFactor = 0.2,
              fillOpacity = 0.5,
  label = paste("Pct membership in civic groups is:",(md2$pct_anymembershp_zip), "Income is",(md2$income_per_capita), "in", (md2$NAME.y), "zip code", (md2$zip))) %>%
 addControl(
    html = '<div id="legend" style="background-color: white; padding: 10px; border-radius: 5px; border: 1px solid #ccc; text-align: center;">
              <p><strong>Pct Membership 2020</strong></p>
              <div style="background-color: gray; height: 20px; width: 30px; display: inline-block;"></div> 0: NA
              <br>
              <div style="background-color: black; height: 20px; width: 30px; display: inline-block;"></div> 30-39%
              <br>
              <div style="background-color: #e75480; height: 20px; width: 30px; display: inline-block;"></div> 40-45%
              <br>
              <div style="background-color: #fa8072; height: 20px; width: 30px; display: inline-block;"></div> 46-50%
              <br>
              <div style="background-color: #ffffbf; height: 20px; width: 30px; display: inline-block;"></div> 51-60%
           </div>',
    position = "bottomright"
  ) %>% 
  addControl(
    html = "<div style='background-color: white; padding: 5px; border-radius: 3px;'><h3>Civil Membership & Income in Maryland</h3></div>",
    position = "topright"
  )

saveWidget(leafMap, "../output/state_member_all_map_jan_28.html")
leafMap
```


### NOTES
# Census Place file - this has problems - missing many values

```{r}
places <- rio::import("https://www2.census.gov/geo/docs/maps-data/data/gazetteer/2020_Gazetteer/2020_gaz_place_24.txt") %>% 
  mutate(geoid1 = as.character(GEOID))



md_places1 <- md %>% 
  left_join(places, by=c("geoid1")) %>% 
  mutate(pct_anymembershp_zip = formattable::percent(pct_anymembershp_zip)) 

```

#Map by county
```{r}
md_named <- read.csv("../output/md_smith_data_12.16.csv") %>% 
  rename(lat = INTPTLAT, long = INTPTLONG)


library(sf)

md_named <- st_as_sf(md_named, coords = c("long", "lat"), crs = 4326)

md_named <- st_set_crs(md_named, 4326) %>%
  st_transform(4326)

md_named <- md_named %>%
  st_as_sf(., coords = c("long", "lat")) %>%
  st_buffer(dist = 0.01) # convert points to polygons

#addCircles(data = md_named, radius = 500, fillOpacity = 0.5, ...)
# #Map by Median Income Difference
# md_named<- md_named %>% 
#    sf::st_transform('+proj=longlat +datum=WGS84')
```


#MD counties
md_counties <- rio::import("../data/md_counties_geoids.txt")

md_named <- md %>% 
  inner_join(md_counties, by=c("county"="GEOID")) %>% 
  mutate(pct_anymembershp_zip = formattable::percent(pct_anymembershp_zip)) 
  
  md_named$income_per_capita <- formattable::currency(md_named$income_per_capita, digits =0L)

#write.csv(md_named, "../output/md_smith_data_1.28.csv")




### Analysis of Maryland Data

```{r}
x<- md_named %>% 
  group_by(zip) %>% 
  count () 
sum(x$n)

#per capita income seems way too low
median(md_named$income_per_capita, na.rm = TRUE) #34917.79

median(md_named$pct_anymembershp_zip, na.rm = TRUE) #0.513087
summary(md_named$pct_anymembershp_zip, na.rm = TRUE) 
 #   Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
 # 0.3124  0.4698  0.5131  0.5150  0.5521  0.6842     150 

median(md_named$nbanks_zip, na.rm = TRUE) #7
summary(md_named$nbanks_zip, na.rm = TRUE) 
# Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
#   0.000   3.000   7.000   8.016  11.000  34.000     147  

median(md_named$census_response_rate2020, na.rm = TRUE) # 0.741
summary(md_named$census_response_rate2020, na.rm = TRUE)
 #  Min. 1st Qu.  Median    Mean 3rd Qu.    Max.    NA's 
 # 0.0650  0.6470  0.7410  0.7161  0.8090  0.9230     122 


```


#Revise for Maryland



## Membership & Income

```{r}
#communities with greater than median membership - 46%- but less than median income - $22,790
himember <- smith %>% 
  group_by(neighborhood) %>% 
  select(pct_anymembershp_zip, income_per_capita, tract) %>% 
  arrange(desc(pct_anymembershp_zip)) %>% 
  filter(pct_anymembershp_zip >= 0.455563) %>% 
  filter(income_per_capita <= 22789.76) %>% 
  mutate(pct_anymembershp_zip = formattable::percent(pct_anymembershp_zip)) 
  
  himember$income_per_capita <- formattable::currency(himember$income_per_capita, digits =0L)
  

library(kableExtra)
himember %>%
  rename(Neighborhood = neighborhood, Membership = pct_anymembershp_zip, Per_Capita_Income = income_per_capita, Census_Tract = tract) %>% 
  kbl(caption = "Communities With Greater than Median Membership - 46%- but Less than Median Income - $22,790", font_size = 50, bold=T) %>%
  kable_classic(full_width = T, html_font = "Cambria") %>%
  column_spec(1, bold = T, border_right = T) %>%
  column_spec(2, border_right = T, width = "10em", background = "yellow") %>% 
  column_spec(3, border_right = T, width = "10em") %>% 
  save_kable(file = "../DataOutput/himember.html", self_contained = T)

```
## Census Response + Income
```{r}
#communities in the upper quartile of census response - 65%- but less than median income - $22,654
hicensus <- smith %>% 
  group_by(neighborhood) %>% 
  select(census_response_rate2020, income_per_capita, pct_anymembershp_zip, tract) %>% 
  arrange(desc(census_response_rate2020)) %>% 
  filter(census_response_rate2020 >= 0.6522) %>% 
  filter(income_per_capita <= 22653.77) %>% 
  mutate(census_response_rate2020 = formattable::percent(census_response_rate2020)) %>% 
  mutate(pct_anymembershp_zip = formattable::percent(pct_anymembershp_zip)) 
  
  hicensus$income_per_capita <- formattable::currency(hicensus$income_per_capita, digits =0L)

  
    hicensus %>%
  rename(Neighborhood = neighborhood, Census_Response = census_response_rate2020, Per_Capita_Income = income_per_capita, Census_Tract = tract, Membership = pct_anymembershp_zip) %>% 
  kbl(caption = "Communities in the Upper Quartile of Census Response - 65%- but Less than Median Income - $22,654", font_size = 50, bold=T) %>%
  kable_classic(full_width = T, html_font = "Cambria") %>%
  column_spec(1, bold = T, border_right = T) %>%
  column_spec(2, border_right = T, width = "10em", background = "yellow") %>% 
  column_spec(3, border_right = T, width = "10em") %>% 
  column_spec(4, border_right = T, width = "10em") %>%     
  save_kable(file = "../DataOutput/hicensus.html", self_contained = T)

```

```{r}

#communities in the upper quartile of census response - 65%- and upper quartile of membership - 46%
smith %>% 
  group_by(neighborhood) %>% 
  select(census_response_rate2020, income_per_capita, pct_anymembershp_zip, nbanks_zip) %>% 
  arrange(desc(census_response_rate2020)) %>% 
  filter(census_response_rate2020 >= 0.6522) %>% 
    filter(pct_anymembershp_zip >= 0.4613) %>% 
  arrange((income_per_capita))


```

